---
title: "FRM_Midterm_Project_1"
output: html_document
---

```{r}
rm(list = ls())
```


```{r}
#library("forecast")
#library("tseries")
library("gridExtra")
library("ggplot2")
library("FKF")
#library("zoo")
library(readxl)
library(data.table)
library(R.matlab)
library(MTS)
library(car)

```




DATA
```{r}

GEdata <- read.csv("GEData.txt",sep= ",", header = TRUE)
UKdata <- read.csv("UKData.txt", sep = ",", header=TRUE)
USdata <- read.csv("USData.txt", sep = ",", header= TRUE)

```


```{r}

#Germany: BMW.DE , DBK.DE
#United kingdom: RSA.L , LRE.L
#USA : COST, FDX, IBM, XOM, DIS, APPL





GE_list <- c("BMW.DE", "DBK.DE")
UK_list <- c("RSA.L", "LRE.L")
US_list <- c("COST","FDX","IBM","XOM","DIS","AAPL")
FX_names <- c("EURUSD.X", "GBPUSD.X")
Indices <- c("X.GDAXI", "X.FTSE", "X.DJI")

GE_data <- subset.data.frame(GEdata, select = GE_list)
UK_data <- subset.data.frame(UKdata, select = UK_list)
US_data <- subset.data.frame(USdata, select = US_list)

FX_data_1 <- subset.data.frame(GEdata, select = FX_names[1])
FX_data_2 <- subset.data.frame(UKdata, select = FX_names[2])

Index_GE <- subset.data.frame(GEdata, select = Indices[1])
Index_UK <- subset.data.frame(UKdata, select = Indices[2])
Index_US <- subset.data.frame(USdata, select = Indices[3])
Index_rates <- cbind(Index_US, Index_GE, Index_UK)



```



Dates
```{r}
dates <- fread("USdata.txt", select = c("Year","Month","Day"))
GEdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
UKdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
USdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

date <- USdata$date
```

Merging data
```{r}
Stocks <- cbind(date,US_data,GE_data,UK_data)
FX_rates <- cbind(date,FX_data_1,FX_data_2)
#Index_rates <- cbind(date,Index_GE, Index_UK)
stock_names <- colnames(Stocks)
```



Converting GE and UK to USD

```{r}

Stocks_converted <- Stocks
Stocks_converted[,8:9] <- Stocks[,8:9]*FX_rates$EURUSD.X
Stocks_converted[,10:11] <- Stocks[,10:11]*FX_rates$GBPUSD.X

```

```{r}
Stocks_converted
```



```{r}

plot(Stocks_converted$date, Stocks_converted$COST/Stocks_converted$COST[1], col = 1, lty = 1, type = "l", ylim=c(0,21), xlab = "Date", ylab="Stock prices in USD")
for (i in 3:11){
  lines(Stocks_converted$date, Stocks_converted[,i]/Stocks_converted[,i][1], col = i, lty = 1, type = "l")
}

legend("topleft", legend = c(stock_names[2:11]), col = c(1:10), lty = c(1,1,1,1,1,1,1,1,1,1), cex = 0.5)

```





X.GDAXI = Germany index
X.FTSE = UK index
X.DJI = US index


Comparing the performance of stocks compared to the corresponding index in USD
```{r}

plot(Stocks_converted$date, (Stocks_converted$BMW.DE/Stocks_converted$BMW.DE[1]), col = 1, lty = 1, type="l", ylim=c(0,5), 
     main = "Germany",
     xlab = "Dates",
     ylab = "Stock performance")
lines(Stocks_converted$date, (Stocks_converted$DBK.DE/Stocks_converted$DBK.DE[1]), col = 2, lty = 1, type="l")
lines(Stocks_converted$date, (Index_rates$X.GDAXI*FX_rates$EURUSD.X/(Index_rates$X.GDAXI[1]*FX_rates$EURUSD.X[1])), col = 3, lty = 1, type = "l")
legend("topleft", legend = c("BMW", "DBK", "GDAXI"), col = c(1,2,3), lty =c(1,1,1),cex=0.8)


```

Plot of UK stocks performance
```{r}
plot(Stocks_converted$date, (Stocks_converted$RSA.L/Stocks_converted$RSA.L[1]), 
     col = 1, 
     lty = 1, 
     type = "l",
     xlab = "Dates",
     ylab = "Stock performance",
     main = "UK",
     ylim = c(0,2.5))
lines(Stocks_converted$date, (Stocks_converted$LRE.L/Stocks_converted$LRE.L[1]), col = 2, lty = 1, type = "l")
lines(Stocks_converted$date, (Index_rates$X.FTSE*FX_rates$GBPUSD.X)/(Index_rates$X.FTSE[1]*FX_rates$GBPUSD.X[1]), col = 3, lty = 1, type ="l")
legend("topleft", legend = c("RSA.L", "LRE.L", "X.FTSE"), col = c(1,2,3), lty =c(1,1,1), cex = 0.8)




```


Plot of USA stock performance
```{r}

plot(Stocks_converted$date, Stocks_converted$COST/Stocks_converted$COST[1], col = 1, lty = 1, type = "l",
     main = "USA",
     ylab = "Stock performance",
     xlab = "Dates",
     ylim = c(0,21))
lines(Stocks_converted$date, Stocks_converted$FDX/Stocks_converted$FDX[1], col = 2, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$IBM/Stocks_converted$IBM[1], col = 3, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$XOM/Stocks_converted$XOM[1], col = 4, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$DIS/Stocks_converted$DIS[1], col = 5, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$AAPL/Stocks_converted$AAPL[1], col = 6, lty = 1, type = "l")
lines(Stocks_converted$date, USdata$X.DJI/USdata$X.DJI[1], col = 7, lty = 1, type = "l")
legend("topleft", legend = c("COST", "FDX", "IBM", "XOM", "DIS","AAPL","X.DJI"), col = c(1,2,3,4,5,6,7), lty =c(1,1,1,1,1,1,1), cex = 0.8)



```



```{r}

returnCalc <- function(x){
  diff(x)/x[-length(x)]
}
```

Calulating daily returns for the stocks
```{r}

n <- nrow(Stocks)
Stock_returns <- matrix(NA,nrow=n-1, ncol = 11)


for (i in 2:11){
  Stock_returns[,i] <- returnCalc(Stocks_converted[,i])
}

#### Matching stocks with corresponding stock names
Stock_returns <- data.frame(Stock_returns)
Stock_returns[,1] <- Stocks$date[2:n]
names(Stock_returns) <- stock_names

```

Computing average yearly returns 
```{r}

Stock_hist_ave <- matrix(NA, nrow = 1, ncol = 10)

for(i in 2:11){
  #Daily average return
  Stock_hist_ave[1,i-1] <- sum(Stock_returns[i])/(n-1)
  #Daily average return converted to average yearly return
  Stock_hist_ave[1,i-1] <- (1+Stock_hist_ave[1,i-1])**252
  # Average yearly return -1 to get the %
  Stock_hist_ave[1,i-1] <- Stock_hist_ave[1,i-1]-1
}

Stock_hist_ave <- data.frame(Stock_hist_ave)
names(Stock_hist_ave) <- stock_names[2:11]
Stock_hist_ave

```


Annualizing the standard deviation
```{r}

Stock_annualized_SD <- matrix(NA,nrow=1,ncol = 10)
for(i in 2:11){
  Stock_annualized_SD[1,i-1] <- sd(Stock_returns[,i])
  Stock_annualized_SD[1,i-1] <- Stock_annualized_SD[1,i-1]*sqrt(252)
}

Stock_annualized_SD <- data.frame(Stock_annualized_SD)
names(Stock_annualized_SD) <- stock_names[2:11]

Stock_annualized_SD

```


#Calculating the correlation between stock returns - covariancee and daily standard deviation
```{r}
Stocks_cor <- cor(Stock_returns[,2:11])
Stocks_cov <- cov(Stock_returns[,2:11])
Stocks_daily_SD <- sqrt(diag(Stocks_cov))

Stocks_cov

```



Conversion of foreign stocks to USD-value and Model Based Approach using equal weights for
determining the covariance matrix - Similar to Question 1 from Week 3

First thing is to create the scenarios with N as 2686

```{r}
n <- nrow(Stocks)
Scenarios <- matrix(NA, nrow = n-1, ncol = 11)

for(i in 2:11){
  for(j in 1:nrow(Scenarios)){
    Scenarios[j,1] <- j
    Scenarios[j,i] <- (Stocks_converted[n,i]*Stocks_converted[j+1,i])/Stocks_converted[j,i]
  }
}

Scenarios <- data.frame(Scenarios)
names(Scenarios) <- c("Scenarios",stock_names[2:11])


```


Calculating the portfolio value for each Scenario
```{r}

#portfolio_loss <- matrix(NA, nrow = nrow(scenarios), ncol = 1)
#portfolio_value <- matrix(NA, nrow = nrow(scenarios), ncol = 1)

weights = seq(10**6, 10**6, length=10)

Scenarios$Portfolio_value <- NA

index_values <- matrix(NA, nrow=nrow(Scenarios), ncol = 10)
for (i in 2:11){
  for (j in 1:nrow(Scenarios)){
  index_values[j,i-1] <- Scenarios[j,i]*t(weights[i-1])/Stocks_converted[n,i]
  }
}


for(j in 1:nrow(Scenarios)){
  Scenarios$Portfolio_value[j] <- rowSums(index_values)[j]
}


Scenarios$Portfolio_loss <- NA
for(j in 1:nrow(Scenarios)){
    Scenarios$Portfolio_loss[j] <- sum(weights)- Scenarios$Portfolio_value[j]
}



```




Histogram of losses
```{r}
hist(Scenarios$Portfolio_loss,  col = "blue", breaks=100)




```


Sorting the losses in decreasing order
```{r}


Scenarios_sorted_loss <-Scenarios[order(Scenarios$Portfolio_loss, decreasing=TRUE),]


Scenarios_sorted_loss


#View(Scenarios_sorted_loss)

```


Finding corresponding 1% & 5% VaR value for the losses
```{r}

percentile_99 <- n*(1-0.99)
percentile_95 <- n*(1-0.95)

cat("For 99% VaR/ES use value:",percentile_99 )
cat(sep="\n\n")
cat("For 95% VaR/ES use value:",percentile_95 )




```

Interpolating to find the VaR value

```{r}

portfolio_VAR_99 <- approx(x=c(26,27), y = c(Scenarios_sorted_loss[26,13],Scenarios_sorted_loss[27,13]), xout = 26.86)
scenario_VAR_99 <- approx(x=c(26,27), y = c(Scenarios_sorted_loss[26,1],Scenarios_sorted_loss[27,1]), xout = 26.86)



portfolio_VAR_95 <- approx(x=c(134,135), y = c(Scenarios_sorted_loss[134,13],Scenarios_sorted_loss[135,13]), xout = 134.3)

scenario_VAR_99 <- approx(x=c(134,135), y = c(Scenarios_sorted_loss[134,1],Scenarios_sorted_loss[135,1]), xout = 134.3)


portfolio_VAR_99<- matrix(unlist(portfolio_VAR_99), ncol = 2, byrow = FALSE)
portfolio_VAR_95 <- matrix(unlist(portfolio_VAR_95), ncol = 2, byrow = TRUE)

portfolio_ES_99 <- mean(Scenarios_sorted_loss[1:26,13])
portfolio_ES_95 <- mean(Scenarios_sorted_loss[1:134,13])





```


Printing out results for VaR and ES for both 99% and 95%
```{r}

cat("Historical Simulation with n = ",nrow(Stocks), "and Scenarios: ", nrow(Scenarios))
cat(sep="\n\n" )
cat("Interpolate on values: 26 and 27 for 99%")
cat(sep="\n\n" )
cat("Simple Approach  99% VAR: ", portfolio_VAR_99[2])
cat(sep="\n\n")
cat("Simple Approach 99% ES:", portfolio_ES_99)
cat(sep="\n\n")
cat("Interpolate on values: 134 and 135 for 95%")
cat(sep="\n\n" )
cat("Simple Approach  95% VAR: ", portfolio_VAR_95[2])
cat(sep="\n\n")
cat("Simple Approach 95% ES:", portfolio_ES_95)
```

Conversion of foreign stocks to USD-value and using Historical Simulation with probability
weighting
```{r}



#### Create lambda for the weights
lambda = 0.995

Scenarios$Weight <- NA

for(j in 1:nrow(Scenarios)){
  Scenarios$Weight[j] <- (lambda**(nrow(Scenarios)-j))*(1-lambda)/(1-(lambda**nrow(Scenarios)))
}

```


Scenarios with Weights - sorted
```{r}


Scenarios_sorted_loss_weights <- Scenarios[order(Scenarios$Portfolio_loss, decreasing=TRUE),]

Scenarios_sorted_loss_weights

```



Cumulative Weights
```{r}

Scenarios_sorted_loss_weights$CumWeight <- NA

Scenarios_sorted_loss_weights$CumWeight[1] <- Scenarios_sorted_loss_weights$Weight[1]

for(j in 2:nrow(Scenarios)){
  Scenarios_sorted_loss_weights$CumWeight[j] <- Scenarios_sorted_loss_weights$CumWeight[j-1]+Scenarios_sorted_loss_weights$Weight[j]
}


```


Finding the cumulative weight corresponding to 0.01 and 0.05
```{r}
tol <- sqrt(.Machine$double.eps)

x <- Scenarios_sorted_loss_weights[,15]

y_99 <- which(abs(x-0.01)==min(abs(x-0.01)))
y_95 <- which(abs(x-0.05)==min(abs(x-0.05)))
cat("The index of in the dataframe is: ", y_99)
cat(sep="\n\n")
cat("The index of in the dataframe is: ", y_95)


```
Finding the VaR and ES with Weights
```{r}
VAR_99_Weights <- Scenarios_sorted_loss_weights$Portfolio_loss[10]
VAR_95_Weights <- Scenarios_sorted_loss_weights$Portfolio_loss[62]

ES_99_Weights <- mean(Scenarios_sorted_loss_weights$Portfolio_loss[1:10])
ES_95_Weights <- mean(Scenarios_sorted_loss_weights$Portfolio_loss[1:62])

```


```{r}
cat("Historical Simulation with probability weighting and with n = ",nrow(Stocks), "and Scenarios: ", nrow(Scenarios))
cat(sep="\n\n" )
cat("Scenario for 99%: 10")
cat(sep="\n\n" )
cat("Weighted Approach  99% VAR: ", VAR_99_Weights)
cat(sep="\n\n")
cat("Weighted Approach 99% ES:", ES_99_Weights)
cat(sep="\n\n")
cat("Scenario for 95%: 62")
cat(sep="\n\n" )
cat("Weighted Approach  95% VAR: ", VAR_95_Weights)
cat(sep="\n\n")
cat("Weighted Approach 95% ES:", ES_95_Weights)
```

Backtesting Historical Simulation with equal weights

```{r}

window = 252



```

