---
title: "FRM_Midterm_Project_1"
output: html_document
---

```{r}
rm(list = ls())
```


```{r}
#library("forecast")
#library("tseries")
library("gridExtra")
library("ggplot2")
library("FKF")
#library("zoo")
library(readxl)
library(data.table)
library(R.matlab)
library(MTS)
library(car)

```




DATA
```{r}

GEdata <- read.csv("GEData.txt",sep= ",", header = TRUE)
UKdata <- read.csv("UKData.txt", sep = ",", header=TRUE)
USdata <- read.csv("USData.txt", sep = ",", header= TRUE)

```


```{r}

#Germany: BMW.DE , DBK.DE
#United kingdom: RSA.L , LRE.L
#USA : COST, FDX, IBM, XOM, DIS, APPL





GE_list <- c("BMW.DE", "DBK.DE")
UK_list <- c("RSA.L", "LRE.L")
US_list <- c("COST","FDX","IBM","XOM","DIS","AAPL")
FX_names <- c("EURUSD.X", "GBPUSD.X")
Indices <- c("X.GDAXI", "X.FTSE", "X.DJI")

GE_data <- subset.data.frame(GEdata, select = GE_list)
UK_data <- subset.data.frame(UKdata, select = UK_list)
US_data <- subset.data.frame(USdata, select = US_list)

FX_data_1 <- subset.data.frame(GEdata, select = FX_names[1])
FX_data_2 <- subset.data.frame(UKdata, select = FX_names[2])

Index_GE <- subset.data.frame(GEdata, select = Indices[1])
Index_UK <- subset.data.frame(UKdata, select = Indices[2])
Index_US <- subset.data.frame(USdata, select = Indices[3])
Index_rates <- cbind(Index_US, Index_GE, Index_UK)



```



Dates
```{r}
dates <- fread("USdata.txt", select = c("Year","Month","Day"))
GEdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
UKdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
USdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

date <- USdata$date
```

Merging data
```{r}
Stocks <- cbind(date,US_data,GE_data,UK_data)
FX_rates <- cbind(date,FX_data_1,FX_data_2)
#Index_rates <- cbind(date,Index_GE, Index_UK)
stock_names <- colnames(Stocks)
```



Converting GE and UK to USD

```{r}

Stocks_converted <- Stocks
Stocks_converted[,8:9] <- Stocks[,8:9]*FX_rates$EURUSD.X
Stocks_converted[,10:11] <- Stocks[,10:11]*FX_rates$GBPUSD.X

```

```{r}
Stocks_converted
```



```{r}

plot(Stocks_converted$date, Stocks_converted$COST/Stocks_converted$COST[1], col = 1, lty = 1, type = "l", ylim=c(0,21), xlab = "Date", ylab="Stock prices in USD")
for (i in 3:11){
  lines(Stocks_converted$date, Stocks_converted[,i]/Stocks_converted[,i][1], col = i, lty = 1, type = "l")
}

legend("topleft", legend = c(stock_names[2:11]), col = c(1:10), lty = c(1,1,1,1,1,1,1,1,1,1), cex = 0.5)

```





X.GDAXI = Germany index
X.FTSE = UK index
X.DJI = US index


Comparing the performance of stocks compared to the corresponding index in USD
```{r}

plot(Stocks_converted$date, (Stocks_converted$BMW.DE/Stocks_converted$BMW.DE[1]), col = 1, lty = 1, type="l", ylim=c(0,5), 
     main = "Germany",
     xlab = "Dates",
     ylab = "Stock performance")
lines(Stocks_converted$date, (Stocks_converted$DBK.DE/Stocks_converted$DBK.DE[1]), col = 2, lty = 1, type="l")
lines(Stocks_converted$date, (Index_rates$X.GDAXI*FX_rates$EURUSD.X/(Index_rates$X.GDAXI[1]*FX_rates$EURUSD.X[1])), col = 3, lty = 1, type = "l")
legend("topleft", legend = c("BMW", "DBK", "GDAXI"), col = c(1,2,3), lty =c(1,1,1),cex=0.8)


```

Plot of UK stocks performance
```{r}
plot(Stocks_converted$date, (Stocks_converted$RSA.L/Stocks_converted$RSA.L[1]), 
     col = 1, 
     lty = 1, 
     type = "l",
     xlab = "Dates",
     ylab = "Stock performance",
     main = "UK",
     ylim = c(0,2.5))
lines(Stocks_converted$date, (Stocks_converted$LRE.L/Stocks_converted$LRE.L[1]), col = 2, lty = 1, type = "l")
lines(Stocks_converted$date, (Index_rates$X.FTSE*FX_rates$GBPUSD.X)/(Index_rates$X.FTSE[1]*FX_rates$GBPUSD.X[1]), col = 3, lty = 1, type ="l")
legend("topleft", legend = c("RSA.L", "LRE.L", "X.FTSE"), col = c(1,2,3), lty =c(1,1,1), cex = 0.8)




```


Plot of USA stock performance
```{r}

plot(Stocks_converted$date, Stocks_converted$COST/Stocks_converted$COST[1], col = 1, lty = 1, type = "l",
     main = "USA",
     ylab = "Stock performance",
     xlab = "Dates",
     ylim = c(0,21))
lines(Stocks_converted$date, Stocks_converted$FDX/Stocks_converted$FDX[1], col = 2, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$IBM/Stocks_converted$IBM[1], col = 3, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$XOM/Stocks_converted$XOM[1], col = 4, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$DIS/Stocks_converted$DIS[1], col = 5, lty = 1, type = "l")
lines(Stocks_converted$date, Stocks_converted$AAPL/Stocks_converted$AAPL[1], col = 6, lty = 1, type = "l")
lines(Stocks_converted$date, USdata$X.DJI/USdata$X.DJI[1], col = 7, lty = 1, type = "l")
legend("topleft", legend = c("COST", "FDX", "IBM", "XOM", "DIS","AAPL","X.DJI"), col = c(1,2,3,4,5,6,7), lty =c(1,1,1,1,1,1,1), cex = 0.8)



```



```{r}

returnCalc <- function(x){
  diff(x)/x[-length(x)]
}
```





Calulating daily returns for the stocks
```{r}

n <- nrow(Stocks)
Stock_returns <- matrix(NA,nrow=n-1, ncol = 11)

for (i in 2:11){
  Stock_returns[,i] <- returnCalc(Stocks_converted[,i])
}

#### Matching stocks with corresponding stock names
Stock_returns <- data.frame(Stock_returns)
Stock_returns[,1] <- Stocks$date[2:n]
names(Stock_returns) <- stock_names


```

Computing average yearly returns 
```{r}

Stock_hist_ave <- matrix(NA, nrow = 1, ncol = 10)

for(i in 2:11){
  #Daily average return
  Stock_hist_ave[1,i-1] <- sum(Stock_returns[i])/(n-1)
  #Daily average return converted to average yearly return
  Stock_hist_ave[1,i-1] <- (1+Stock_hist_ave[1,i-1])**252
  # Average yearly return -1 to get the %
  Stock_hist_ave[1,i-1] <- Stock_hist_ave[1,i-1]-1
}

Stock_hist_ave <- data.frame(Stock_hist_ave)
names(Stock_hist_ave) <- stock_names[2:11]
Stock_hist_ave



```


Annualizing the standard deviation
```{r}

Stock_annualized_SD <- matrix(NA,nrow=1,ncol = 10)
for(i in 2:11){
  Stock_annualized_SD[1,i-1] <- sd(Stock_returns[,i])
  Stock_annualized_SD[1,i-1] <- Stock_annualized_SD[1,i-1]*sqrt(252)
}

Stock_annualized_SD <- data.frame(Stock_annualized_SD)
names(Stock_annualized_SD) <- stock_names[2:11]

Stock_annualized_SD

```


#Calculating the correlation between stock returns - covariancee and daily standard deviation
```{r}
Stocks_cor <- cor(Stock_returns[,2:11])
Stocks_cov <- cov(Stock_returns[,2:11])
Stocks_daily_SD <- sqrt(diag(Stocks_cov))

Stocks_cov

```


Conversion of foreign stocks to USD-value and Model Based Approach using equal weights for
determining the covariance matrix
```{r}
weights = seq(10**6,10**6, length = 10)

var_equal_weights <- t(weights) %*% Stocks_cov %*%weights

sd_equal_weights <- sqrt(var_equal_weights)

```

Calculating 99% and 95% VaR and ES 
```{r}
Z_99 <- qnorm(0.99,0,1)
Z_95 <- qnorm(0.95,0,1)

ONEday_VAR_99 <- Z_99*sd_equal_weights
ONEday_ES_99 <- sd_equal_weights*exp(-(Z_99^2)/2)/(sqrt(2*pi)*0.01)

ONEday_VAR_95 <- Z_95*sd_equal_weights
ONEday_ES_95 <- sd_equal_weights*exp(-(Z_95^2)/2)/(sqrt(2*pi)*0.05)


```

Printing out results
```{r}

cat("Model Based Approach using equal weights for determining Covariance Matrix")
cat(sep="\n\n")
cat("Equal weight Portfolio for:", nrow(Stocks), "data points")
cat(sep="\n\n")
cat("One day 99% VaR:", ONEday_VAR_99, "and ES:", ONEday_ES_99)
cat(sep="\n\n")
cat("One day 95% VaR:", ONEday_VAR_95, "and ES:", ONEday_ES_95)
```



Conversion of foreign stocks to USD-value and Model Based Approach using EWMA weights for
determining the covariance matrix (with N+1?)

Population variance not sample variance
```{r}
varp <- function(x) mean((x-mean(x))^2)
```


Beginning
```{r}
lambda <- 0.94

#Variance_EWMA <- matrix(NA, nrow = nrow(Stock_returns)+1, ncol = ncol(Stock_returns[2:11]))

HistoricalReturns <- data.matrix(Stock_returns[,2:11])
Cov_Matrix = cov(Stock_returns[,2:11]) 
Variances_EWMA <- Cov_Matrix


#Calclating the covariance-variance matrix
for (i in 2:length(Stocks$date)){
  Variances_EWMA <- lambda*Variances_EWMA + (1-lambda)*HistoricalReturns[i-1,]%*%t(HistoricalReturns[i-1,]);
}





```



```{r}
FinalVols_1 <- 100*diag(Cov_Matrix)**0.5

FinalVols_2 <- 100*diag(Variances_EWMA)**0.5

SD <-diag(Variances_EWMA)**0.5

PF_variance_EWMA <- t(weights)%*%Variances_EWMA%*%weights

VAR99_EWMA <- qnorm(0.99)*sqrt(PF_variance_EWMA)

ES99_EWMA <- sqrt(PF_variance_EWMA)*exp(-qnorm(0.99)^2/2)/sqrt(2*pi)/0.01


VAR_95_EWMA <- qnorm(0.95)*sqrt(PF_variance_EWMA)
ES95_EWMA <- sqrt(PF_variance_EWMA)*exp(-qnorm(0.95)^2/2)/sqrt(2*pi)/0.05

```

Printing out results
```{r}

cat("Model Based Approach using EWMA weights for determining the covariance matrix")
cat(sep="\n\n")
cat("MBA - EWMA Portfolio for:", nrow(Stocks), "data points")
cat(sep="\n\n")
cat("One day 99% VaR:", VAR99_EWMA, "and ES:", ES99_EWMA)
cat(sep="\n\n")
cat("One day 95% VaR:", VAR_95_EWMA, "and ES:", ES95_EWMA)
```

Backtesting on MBA using Equal weights for determining the covariance matrix.

```{r}
window = 252



```

