---
title: "FRM_Midterm_Project_1"
output: html_document
---

```{r}
rm(list = ls())
```


```{r}
#library("forecast")
#library("tseries")
library("gridExtra")
library("ggplot2")
library("FKF")
#library("zoo")
library(readxl)
library(data.table)
library(R.matlab)
library(MTS)
library(car)

```




DATA
```{r}

GEdata <- read.csv("GEData.txt",sep= ",", header = TRUE)
UKdata <- read.csv("UKData.txt", sep = ",", header=TRUE)
USdata <- read.csv("USData.txt", sep = ",", header= TRUE)

```


```{r}

#Germany: BMW.DE , DBK.DE
#United kingdom: RSA.L , LRE.L
#USA : COST, FDX, IBM, XOM, DIS, APPL





GE_list <- c("BMW.DE", "DBK.DE")
UK_list <- c("RSA.L", "LRE.L")
US_list <- c("COST","FDX","IBM","XOM","DIS","AAPL")
FX_names <- c("EURUSD.X", "GBPUSD.X")
Indices <- c("X.GDAXI", "X.FTSE", "X.DJI")

GE_data <- subset.data.frame(GEdata, select = GE_list)
UK_data <- subset.data.frame(UKdata, select = UK_list)
US_data <- subset.data.frame(USdata, select = US_list)

FX_data_1 <- subset.data.frame(GEdata, select = FX_names[1])
FX_data_2 <- subset.data.frame(UKdata, select = FX_names[2])

Index_GE <- subset.data.frame(GEdata, select = Indices[1])
Index_UK <- subset.data.frame(UKdata, select = Indices[2])
Index_US <- subset.data.frame(USdata, select = Indices[3])
Index_rates <- cbind(Index_US, Index_GE, Index_UK)



```



Dates
```{r}
dates <- fread("USdata.txt", select = c("Year","Month","Day"))
GEdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
UKdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")
USdata$date <- as.Date(with(dates, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

date <- USdata$date
```

Merging data
```{r}
Stocks <- cbind(date,US_data,GE_data,UK_data)
FX_rates <- cbind(date,FX_data_1,FX_data_2)
#Index_rates <- cbind(date,Index_GE, Index_UK)
stock_names <- colnames(Stocks)
```



Converting GE and UK to USD

```{r}

Stocks_converted <- Stocks
Stocks_converted[,8:9] <- Stocks[,8:9]*FX_rates$EURUSD.X
Stocks_converted[,10:11] <- Stocks[,10:11]*FX_rates$GBPUSD.X

```


```{r}

returnCalc <- function(x){
  diff(x)/x[-length(x)]
}
```


Calulating daily returns for the stocks
```{r}

n <- nrow(Stocks)
Stock_returns <- matrix(NA,nrow=n-1, ncol = 11)
Stock_returns_converted <- matrix(NA,nrow=n-1, ncol = 11)

for (i in 2:11){
  Stock_returns_converted[,i] <- returnCalc(Stocks_converted[,i])
}

for (i in 2:11){
  Stock_returns[,i] <- returnCalc(Stocks[,i])
}

#### Matching stocks with corresponding stock names
Stock_returns <- data.frame(Stock_returns)
Stock_returns[,1] <- Stocks$date[2:n]
names(Stock_returns) <- stock_names


Stock_returns_converted <- data.frame(Stock_returns_converted)
Stock_returns_converted[,1] <- Stocks$date[2:n]
names(Stock_returns_converted) <- stock_names

```

```{r}
Stocks
```


Computing average yearly returns 
```{r}

Stock_hist_ave <- matrix(NA, nrow = 1, ncol = 10)

for(i in 2:11){
  #Daily average return
  Stock_hist_ave[1,i-1] <- sum(Stock_returns[i])/(n-1)
  #Daily average return converted to average yearly return
  Stock_hist_ave[1,i-1] <- (1+Stock_hist_ave[1,i-1])**252
  # Average yearly return -1 to get the %
  Stock_hist_ave[1,i-1] <- Stock_hist_ave[1,i-1]-1
}

Stock_hist_ave <- data.frame(Stock_hist_ave)
names(Stock_hist_ave) <- stock_names[2:11]
Stock_hist_ave



```


Annualizing the standard deviation
```{r}

Stock_annualized_SD <- matrix(NA,nrow=1,ncol = 10)
for(i in 2:11){
  Stock_annualized_SD[1,i-1] <- sd(Stock_returns[,i])
  Stock_annualized_SD[1,i-1] <- Stock_annualized_SD[1,i-1]*sqrt(252)
}

Stock_annualized_SD <- data.frame(Stock_annualized_SD)
names(Stock_annualized_SD) <- stock_names[2:11]

Stock_annualized_SD

```


#Calculating the correlation between stock returns - covariancee and daily standard deviation
```{r}
Stocks_cor <- cor(Stock_returns[,2:11])
Stocks_cov <- cov(Stock_returns[,2:11])
Stocks_daily_SD <- sqrt(diag(Stocks_cov))

Stocks_cov

```




```{r}
head(FX_rates)
```


Mapping of portfolio into stock returns and FX returns (total of 10+2 risk factors) and using Model
Based Approach


Begin by calculating the returns of the exchange rates and merging datasets
```{r}
FX_returns <- matrix(NA,nrow=n-1, ncol = 2)



for (i in 2:3){
  FX_returns[,i-1] <- returnCalc(FX_rates[,i])
}

Stock_returns_FX <- Stock_returns

Stock_returns_FX[,12] <- FX_returns[,1]
Stock_returns_FX[,13] <-FX_returns[,2]

names(Stock_returns_FX)[names(Stock_returns_FX)=="V12"] <- "EURUSD"
names(Stock_returns_FX)[names(Stock_returns_FX)=="V13"] <- "GBPUSD"


```

```{r}
head(Stock_returns_FX)
```


Calculating index returns
```{r}
Index_returns <- matrix(NA,nrow=n-1, ncol = 3)

for (i in 1:3){
  Index_returns[,i] <- returnCalc(Index_rates[,i])
}



```

Converting the structure into matrix so the linear model can read it
```{r}
stocks_EUR <- as.matrix(Stock_returns[,8:9])
index_EUR <- as.matrix(Index_returns[,2])
fx_EUR <- as.matrix(FX_returns[,1])
stocks_GBP <- as.matrix(Stock_returns[,10:11])
fx_GBP <- as.matrix(FX_returns[,2])



```

Creating linear model for the stock returns over market returns
```{r}
fit_EUR_fx <- lm(stocks_EUR~fx_EUR)
fit_GBP_fx <- lm(stocks_GBP~fx_GBP)
options(digits=5)

```

Betas for each model
```{r}
beta_EUR_fx <- fit_EUR_fx$coef[2,]
beta_GBP_fx <- fit_GBP_fx$coef[2,]
```



Creating a new weight portfolio and constructing the covariance matrix
```{r}

weights_FX <- c(10**6, 10**6, 10**6, 10**6, 10**6, 10**6, 10**6, 10**6, 10**6, 10**6, 2*10**6, 2*10**6)


weights_FX
Exposures_FX <- matrix(NA, nrow=12, ncol=1)

for(i in 1:10){
  Exposures_FX[i,1] <- 10**6
}
Exposures_FX[11,1] <- 2*(10**6)*sum(beta_EUR_fx)
Exposures_FX[12,1] <- 2*(10**6)*sum(beta_GBP_fx)


Cov_with_FX <- cov(Stock_returns_FX[,2:13])




```

Variance and standard deviation of the portfolio with FX rates
```{r}
portfolio_Var_FX <- t(Exposures_FX)%*%Cov_with_FX%*%Exposures_FX

portfolio_Std_FX <- sqrt(portfolio_Var_FX)
```


Calculating one day 99-VAR/ES and 95-VAR/ES
```{r}
Z_99 <- qnorm(0.99,0,1)
Z_95 <- qnorm(0.95,0,1)

VAR_99_FX <- Z_99*portfolio_Std_FX
ES_99_FX <- portfolio_Std_FX*exp(-(Z_99^2)/2)/(sqrt(2*pi)*0.01)

VAR_95_FX <- Z_95*portfolio_Std_FX
ES_95_FX <- portfolio_Std_FX*exp(-(Z_95^2)/2)/(sqrt(2*pi)*0.05)

```
Printing out results
```{r}
cat("Model Based Approach with:", n, "data points")
cat(sep="\n\n")
cat("Mapped portfolio into Stock returns and FX returns")
cat(sep="\n\n")
cat("(10+2 risk factors)")
cat(sep="\n\n")
cat("One day 99% VaR:", VAR_99_FX, "and ES:", ES_99_FX)
cat(sep="\n\n")
cat("One day 95% VaR:", VAR_95_FX, "and ES:", ES_95_FX)
```
One day 99% VaR: 265681 and ES: 304381.3
One day 95% VaR: 187850.8 and ES: 235572.5


```{r}

265681/258977
304381/296701
187850.8/183111
235572.5/229629


(1.0259-1)*100
```


Mapping of portfolio into market index returns and FX returns(total of 3+2 risk factors) and using
Model Based Approach



Begin by calulcating the return of the Stock indices returns and FX returns and merging them together

```{r}
head(Index_rates)
```


```{r}
head(FX_rates)
```




Converting the structure of the matrices to data.frames and renaming the columns to corresponding values
```{r}


FX_returns <- data.frame(FX_returns)
Index_returns <- data.frame(Index_returns)




names(FX_returns)[names(FX_returns)=="X1"] <- "EURUSD"
names(FX_returns)[names(FX_returns)=="X2"] <- "GBPUSD"
names(Index_returns)[names(Index_returns)=="X1"] <- "X.DJI"
names(Index_returns)[names(Index_returns)=="X2"] <- "X.GDAXI"
names(Index_returns)[names(Index_returns)=="X3"] <- "X.FTSE"

```


Merging Index and FX returns together
```{r}
Index_FX_returns <- cbind(Index_returns, FX_returns)
```

```{r}
head(Index_FX_returns)
```
```{r}
head(FX_returns)
```

```{r}
head(Index_returns)

```




Converting the structure into matrix so the linear model can read it
```{r}
stocks_US <- as.matrix(Stock_returns[,2:7])
index_US <- as.matrix(Index_returns[,1])


stocks_EUR <- as.matrix(Stock_returns[,8:9])
index_EUR <- as.matrix(Index_returns[,2])
fx_EUR <- as.matrix(FX_returns[,1])

stocks_GBP <- as.matrix(Stock_returns[,10:11])
index_GBP <- as.matrix(Index_returns[,3])
fx_GBP <- as.matrix(FX_returns[,2])

```


Creating linear model for the stock returns over market returns
```{r}
fit_US <- lm(stocks_US~index_US)
fit_EUR <-lm(stocks_EUR~index_EUR)
fit_GBP <- lm(stocks_GBP~index_GBP)

fit_EUR_fx <- lm(stocks_EUR~fx_EUR)
fit_GBP_fx <- lm(stocks_GBP~fx_GBP)
options(digits=5)

```

Betas for each model
```{r}
beta_US <- fit_US$coef[2,]
beta_EUR <- fit_EUR$coef[2,]
beta_GBP <- fit_GBP$coef[2,]

beta_EUR_fx <- fit_EUR_fx$coef[2,]
beta_GBP_fx <- fit_GBP_fx$coef[2,]
```


```{r}
Index_returns
```

Covariance matrix from the returns
```{r}
Cov2Factor <- cov(cbind(Index_returns,FX_returns))

```

Constructing new weights portfolio
```{r}
Exposures <- matrix(NA, nrow = 5, ncol = 1)

Exposures[1,1] = (10**6)*sum(beta_US)
Exposures[2,1] = (10**6)*sum(beta_EUR)
Exposures[3,1] = (10**6)*sum(beta_GBP)
Exposures[4,1] = (10**6)*sum(beta_EUR_fx)
Exposures[5,1] = (10**6)*sum(beta_GBP_fx)

Exposures


```



Calculating one day 99-VAR/ES and 95-VAR/ES
```{r}
PF_Var2 <- t(Exposures) %*%Cov2Factor%*%(Exposures)
VaR99_FX_Index <- sqrt(PF_Var2)*Z_99
ES99_FX_Index <- sqrt(PF_Var2)*exp(-(Z_99^2)/2)/(sqrt(2*pi)*0.01)
VaR_95_FX_Index <- sqrt(PF_Var2)*Z_95
ES95_FX_Index <- sqrt(PF_Var2)*exp(-(Z_95^2)/2)/(sqrt(2*pi)*0.05)
```




Printing out 1 day VAR and ES for 99% and 95% using MBA portfolio mapping into index returns and FX returns
```{r}
cat("Model Based Approach with:", n, "data points")
cat(sep="\n\n")
cat("Mapped portfolio into market index returns and FX returns")
cat(sep="\n\n")
cat("(3+2 risk factors)")
cat(sep="\n\n")
cat("One day 99% VaR:", VaR99_FX_Index, "and ES:", ES99_FX_Index)
cat(sep="\n\n")
cat("One day 95% VaR:", VaR_95_FX_Index, "and ES:", ES95_FX_Index)
```


```{r}
# Backtesting


var_95_series_m2 = c();
es_95_series_m2 = c()
var_99_series_m2 = c();
es_99_series_m2 = c();

N = 2685

for (i in 1:252){

  stocks_US_backtesting <- stocks_US[i:(N-252+i),]
  stocks_EUR_backtesting <- stocks_EUR[i:(N-252+i),]
  stocks_GBP_backtesting <- stocks_GBP[i:(N-252+i),]
  
  index_US_backtesting <- index_US[i:(N-252+i),]
  index_EUR_backtesting <- index_EUR[i:(N-252+i),]
  index_GBP_backtesting <- index_GBP[i:(N-252+i),]
  fx_EUR_backtesting <- fx_EUR[i:(N-252+i),]
  fx_GBP_backtesting <- fx_GBP[i:(N-252+i),]

  fit_US <- lm(stocks_US_backtesting~index_US_backtesting)
  fit_EUR <-lm(stocks_EUR_backtesting~index_EUR_backtesting)
  fit_GBP <- lm(stocks_GBP_backtesting~index_GBP_backtesting)
  
  fit_EUR_fx <- lm(stocks_EUR_backtesting~fx_EUR_backtesting)
  fit_GBP_fx <- lm(stocks_GBP_backtesting~fx_GBP_backtesting)
  options(digits=5)
  
  beta_US <- fit_US$coef[2,]
  beta_EUR <- fit_EUR$coef[2,]
  beta_GBP <- fit_GBP$coef[2,]
  
  beta_EUR_fx <- fit_EUR_fx$coef[2,]
  beta_GBP_fx <- fit_GBP_fx$coef[2,]
  
  Cov2Factor <- cov(cbind(Index_returns[i:(N-252+i),],FX_returns[i:(N-252+i),]))
  
  Exposures <- matrix(NA, nrow = 5, ncol = 1)
  
  Exposures[1,1] = (10**6)*sum(beta_US)
  Exposures[2,1] = (10**6)*sum(beta_EUR)
  Exposures[3,1] = (10**6)*sum(beta_GBP)
  Exposures[4,1] = (10**6)*sum(beta_EUR_fx)
  Exposures[5,1] = (10**6)*sum(beta_GBP_fx)
  
  PF_Var2 <- t(Exposures) %*%Cov2Factor%*%(Exposures)
  VaR99_FX_Index <- sqrt(PF_Var2)*Z_99
  ES99_FX_Index <- sqrt(PF_Var2)*exp(-(Z_99^2)/2)/(sqrt(2*pi)*0.01)
  VaR_95_FX_Index <- sqrt(PF_Var2)*Z_95
  ES95_FX_Index <- sqrt(PF_Var2)*exp(-(Z_95^2)/2)/(sqrt(2*pi)*0.05)
  
  var_95_series_m2 = append(var_95_series_m2, VaR_95_FX_Index)
  es_95_series_m2 = append(es_95_series_m2, ES95_FX_Index)
  var_99_series_m2 = append(var_99_series_m2, VaR99_FX_Index)
  es_99_series_m2 = append(es_99_series_m2, ES99_FX_Index)
}



```


```{r}

pnl = rowSums(Stock_returns_converted[2:11] * c(1000000, 1000000, 1000000, 1000000, 1000000, 1000000, 1000000, 1000000, 1000000, 1000000)[col(Stock_returns_converted)])
names(pnl) <- Stock_returns_converted[,1]

dates_bt = tail(Stock_returns_converted$date, 252)

plot(dates_bt, tail(pnl, 252), col = 1, lty = 1, type = "l",
     main = "Backtesting 95% VaR",
     ylab = "PNL",
     xlab = "Dates")
lines(dates_bt, -var_95_series_m2, col = 2, lty = 1, type = "l")
legend("topleft", legend = c("PnL", "VaR 95%"), col = c(1,2), lty =c(1,1), cex = 0.8)
```

```{r}
plot(dates_bt, tail(pnl, 252), col = 1, lty = 1, type = "l",
     main = "Backtesting 99% VaR",
     ylab = "PNL",
     xlab = "Dates")
lines(dates_bt, -var_99_series_m2, col = 2, lty = 1, type = "l")
legend("topleft", legend = c("PnL", "VaR 99%"), col = c(1,2), lty =c(1,1), cex = 0.8)
```
```{r}
# PNL values
pnl_values = tail(pnl, 252)

# number where loss exceeds VaR
n1 = sum(pnl_values + var_95_series_m2 < 0)
n0 = 252-n1
n1

sigma_exp = 0.05
sigma_obs = n1 / 252 # observed proportion of exceedances

test_statistic_LR_uc = ((sigma_exp**n1)*(1-sigma_exp)**n0) / ((sigma_obs**n1)*(1-sigma_obs)**n0)
LR_uc = -2 * log(test_statistic_LR_uc)

# Deny if 10% > 2.71 and at a 5% level, if LR_uc > 3.84
LR_uc
```

```{r}
# PNL values
pnl_values = tail(pnl, 252)

# number where loss exceeds VaR
n1 = sum(pnl_values + var_99_series_m2 < 0)
n0 = 252-n1
n1

sigma_exp = 0.01
sigma_obs = n1 / 252 # observed proportion of exceedances

test_statistic_LR_uc = ((sigma_exp**n1)*(1-sigma_exp)**n0) / ((sigma_obs**n1)*(1-sigma_obs)**n0)
LR_uc = -2 * log(test_statistic_LR_uc)

# Deny if 10% > 2.71 and at a 5% level, if LR_uc > 3.84
LR_uc
```

